Structured Versus Decorated Cospans

### 2020-11-27 cospans.tex
Added a diagram for the associator for horizontal 1-cell composition in Theorem 2.2

\begin{thm}\label{thm:decorated_cospans}
Let $\A$ be a category with finite colimits, made monoidal with a choice of initial object and binary coproducts.  Let $F \maps (\A,+) \to (\Cat,\times)$ be a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a cospan in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $x \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a map of cospans in $\A$:
\begin{equation}\label{eq:mapofdec}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[above] {$i'$} (D)
(C') edge node[above] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\end{equation}
together with a \define{decoration morphism} $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}
\end{thm}

\begin{proof}
Take $\A$ to have finite colimits and let $(F,\phi,\phi_0) \maps (A,+,0) \to (\Cat,\times, 1)$ be a lax monoidal pseudofunctor as defined in \cref{defn:double_category}.  We shall define the category of objects $F\lCsp_0$, the category of arrows $F\lCsp_1$, as well as the four structure functors which tie them together: the source and target functors $S,T \colon F\lCsp_1 \to F\lCsp_0$, the composition functor $\odot \colon F\lCsp_1 \times_{F\lCsp_0} F\lCsp_1 \to F\lCsp_1$ and the unit functor $U \colon F\lCsp_0 \to F\lCsp_1$.

The category of objects $F\lCsp_0$ is simply $\A$, whereas $F$-decorated cospans and maps between them form the category of arrows $F\lCsp_1$. The (vertical) composition of two $F$-decorated cospans maps is 

@@

The associator $\alpha \colon \odot (\odot \times 1) \Rightarrow \odot (1 \times \odot)$ is formed as follows: for three composable horizontal 1-cells
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$M_1=a$};
\node (B) at (1,0) {$m_1$};
\node (C) at (2,0) {$b$};
\node (D) at (1,-0.5) {$x_1 \in F(m_1)$};
\node (E) at (3,0) {$M_2=b$};
\node (F) at (4,0) {$m_2$};
\node (G) at (5,0) {$c$};
\node (H) at (4,-0.5) {$x_2 \in F(m_2)$};
\node (I) at (6,0) {$M_3=c$};
\node (J) at (7,0) {$m_3$};
\node (K) at (8,0) {$d$};
\node (L) at (7,-0.5) {$x_3 \in F(m_3)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(E) edge node[above]{$i'$} (F)
(G) edge node[above]{$o'$} (F)
(I) edge node[above]{$i''$} (J)
(K) edge node[above]{$o''$} (J);
\end{tikzpicture}
\]
the isomorphism $(M_1\odot M_2)\odot M_3\cong M_1\odot (M_2\odot M_3)$ is the isomorphism of cospans 
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a$};
\node (B) at (1.5,0.5) {$(m_1+_{b} m_2)+_{c} m_3$};
\node (C) at (3,0.5) {$d$};
\node (C') at (3,-0.5) {$d$};
\node (D) at (1.5,-0.5) {$m_1+_{b}(m_2 +_{c} m_3)$};
\node (E) at (5.5,0.5) {$(x_1 \odot x_2) \odot x_3 \in F((m_1+_{b} m_2)+_{c} m_3)$};
\node (F) at (5.5,-0.5) {$x_1 \odot (x_2 \odot x_3) \in F(m_1+_{b} (m_2 +_{c} m_3))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$$} (B)
(C) edge node[above]{$$} (B)
(A) edge node[left]{$\id_a$} (A')
(C) edge node[right]{$\id_d$} (C')
(A') edge node {$$} (D)
(C') edge node {$$} (D)
(B) edge node [left] {$\stackrel{\alpha}{\cong}$} (D);
\end{tikzpicture}
\]
together with an isomorphism $F(\alpha)((x_1 \odot x_2) \odot x_3) \cong (x_1 \odot (x_2 \odot x_3))$ between the decorations 
\begin{align*}
\scalebox{.8}{$\one\xrightarrow{x_1\times x_2}F(m_1)\times F(m_2)\xrightarrow{\phi}F(m_1+m_2)\to F(m_1+_bm_2)\xrightarrow{1\times x_3}F(m_1+_bm_2)\times F(m_3)\xrightarrow{\phi}F((m_1+_bm_2)+m_3)$}&\scalebox{.8}{$\to F((m_1+_bm_2)+_cm_3)$} \\
&\scalebox{.8}{$\stackrel{F(\alpha)}{\cong} F(m_1+_b(m_2+_cm_3))$} \\
\scalebox{.8}{$\one\xrightarrow{x_2\times x_3}F(m_2)\times F(m_3)\xrightarrow{\phi}F(m_2+m_3)\to F(m_2+_cm_3)\xrightarrow{x_1\times1}F(m_1)\times F(m_2+_cm_3)\xrightarrow{\phi}F(m_1\times(m_2+_cm_3))$}&\scalebox{.8}{$\to F(m_1+_b(m_2+_cm_3))$}
\end{align*}
which is constructed by pasting structure isomorphisms like above. Similarly the right and left unitors can be constructed, and all data satisfy the axioms of a double category -- essentially breaking down to those of the cospan double category of $\A$ and axioms of the lax monoidal pseudofunctor $F$. For full details of these arguments, see the second author's thesis \cite[Theorem~4.1.1]{CourserThesis}.
\begin{comment}
 The definition of double category may be found in Definition \ref{defn:double_category}. To show that the pseudofunctor $F$ in the theorem statement induces a double category $F\lCsp$, we must first define the category of objects $F\lCsp_0$, the category of arrows $F\lCsp_1$, as well as the four structure functors which tie them together: the source and target functors $S,T \colon F\lCsp_1 \to F\lCsp_0$, the unit functor $U \colon F\lCsp_0 \to F\lCsp_1$ and the composition functor $\odot$ which says how to compose two composable horizontal 1-cells or two composable 2-morphisms horizontally: $$\odot \colon F\lCsp_1 \times_{F\lCsp_0} F\lCsp_1 \to F\lCsp_1.$$

@@

Let $\A$ be a category with finite colimits, made symmetric monoidal with a choice of initial object and binary coproducts.   Let $F \maps (\A,+)\to (\Cat,\times)$ be symmetric lax monoidal pseudofunctor. Then the double category $F\lCsp$ of Theorem \ref{thm:decorated_cospans} is symmetric monoidal, where

@@

\begin{proof}
Monoidal and symmetric monoidal double categories are introduced in Definitions \ref{defn:monoidal_double_category} and \ref{defn:symmetric_monoidal_double_category}. To make $F\lCsp$ into a symmetric monoidal double category, we must show that both the category of objects $F\lCsp_0$ and the category of arrows $F\lCsp_1$ are symmetric monoidal, show that the monoidal unit of $F\lCsp_0$ is preserved by the unit functor $U \maps F\lCsp_0 \to F\lCsp_1$, show that the source and target functors $S,T \colon F\lCsp_1 \to F\lCsp_0$ are strict symmetric monoidal, define the globular isomorphisms $\chi$ and $\mu$ that relate how composition and units interact with the tensor product, and check that the required axioms are satisfied.

First we note that the category of objects $F\lCsp_0=\A$ is symmetric monoidal under chosen binary coproducts in $\A$ and the left and right unitors, associators and braidings are given as natural maps. Next we show that $F\lCsp_1$ is symmetric monoidal. The category of arrows $F\lCsp_1$ has:
\begin{enumerate}
\item{objects as $F$-decorated cospans, which are pairs:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
and}
\item{morphisms as maps of $F$-decorated cospans, which are cospans in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node [above]{$i'$} (D)
(C') edge node [above]{$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a decoration morphism $\tau \maps F(h)(x) \to x'$.
}
\end{enumerate}
First we define the tensor product and monoidal unit of $F\lCsp_1$. Given two objects $M_1$ and $M_2$ of $F\lCsp_1$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$m_1$};
\node (C) at (2,0) {$b_1$};
\node (D) at (1,-0.5) {$x_1 \in F(m_1)$};
\node (E) at (3,0) {$a_2$};
\node (F) at (4,0) {$m_2$};
\node (G) at (5,0) {$b_2$};
\node (H) at (4,-0.5) {$x_2 \in F(m_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(E) edge node[above]{$i_2$} (F)
(G) edge node[above]{$o_2$} (F);
\end{tikzpicture}
\]
their tensor product $M_1 \otimes M_2$ is given by taking the coproducts of the cospans in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1+a_2$};
\node (B) at (1.5,0) {$m_1+m_2$};
\node (C) at (3,0) {$b_1+b_2$};
\node (D) at (4.55,0) {$x_1+x_2 \in F(m_1+m_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B);
\end{tikzpicture}
\]
where the decoration on the apex is obtained using the natural transformation of the symmetric lax monoidal pseudofunctor $F$: $$x_1+x_2 \mapseqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x_1 \times x_2} F(m_1) \times F(m_2) \xrightarrow{\phi_{m_1,m_2}} F(m_1+m_2).$$The monoidal unit $0_{F\lCsp_1}$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$0$};
\node (B) at (1,0) {$0$};
\node (C) at (2,0) {$0$};
\node (D) at (3,0) {$\bot_0 \in F(0)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$!$} (B)
(C) edge node[above]{$!$} (B);
\end{tikzpicture}
\]
where $0$ is the monoidal unit of $\A$ and the element $\bot_0$ is another name for $\phi \maps 1 \to F(0)$, the morphism that arises from the lax monoidal structure of $F \maps (\A,+) \to (\Cat,\times)$. It is also clear that $U(0)=0_{F\lCsp_1}$.

>>>>

\begin{thm}\label{thm:decorated_cospans}
Let $\A$ be a category with finite colimits and $F \maps \A \to \Cat$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}, that is, 
a cospan in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $x \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}, that is, 
a map of cospans in $\A$:
\begin{equation}\label{eq:mapofdec}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[above] {$i'$} (D)
(C') edge node[above] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\end{equation}
together with a \define{decoration morphism} $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}
\end{thm}

\begin{proof}
According to \cref{defn:double_category}, for a lax monoidal pseudofunctor $(F,\phi,\phi_0)$ we will define the category of objects $F\lCsp_0$, the category of arrows $F\lCsp_1$, as well as the four structure functors which tie them together: the source and target functors $S,T \colon F\lCsp_1 \to F\lCsp_0$, the composition functor $\odot \colon F\lCsp_1 \times_{F\lCsp_0} F\lCsp_1 \to F\lCsp_1$ and the unit functor $U \colon F\lCsp_0 \to F\lCsp_1$.

The category of objects $F\lCsp_0$ is simply the category $\A$, whereas $F$-decorated cospans and maps between them form the category of arrows $F\lCsp_1$. The (vertical) composition of two $F$-decorated cospans maps is 

@@

The associator $\alpha \colon \odot (\odot \times 1) \Rightarrow \odot (1 \times \odot)$ is formed as follows: for three composable horizontal 1-cells
\node (A) at (0,0) {$M_1=a$};
\node (E) at (3,0) {$M_2=b$};
\node (I) at (6,0) {$M_3=c$};
the isomorphism $(M_1\odot M_2)\odot M_3\cong M_1\odot (M_2\odot M_3)$ is the isomorphism of cospans 
(B) edge node [left] {$\stackrel{\alpha}{\cong}$} (D);
together with an isomorphism $F(\alpha)((x_1 \odot x_2) \odot x_3) \cong (x_1 \odot (x_2 \odot x_3))$ between the decorations 
\begin{align*}
\scalebox{.8}{$\one\xrightarrow{x_1\times x_2}F(m_1)\times F(m_2)\xrightarrow{\phi}F(m_1+m_2)\to F(m_1+_bm_2)\xrightarrow{1\times x_3}F(m_1+_bm_2)\times F(m_3)\xrightarrow{\phi}F((m_1+_bm_2)+m_3)$}&\scalebox{.8}{$\to F((m_1+_bm_2)+_cm_3)$} \\
&\scalebox{.8}{$\stackrel{F(\alpha)}{\cong} F(m_1+_b(m_2+_cm_3))$} \\
\scalebox{.8}{$\one\xrightarrow{x_2\times x_3}F(m_2)\times F(m_3)\xrightarrow{\phi}F(m_2+m_3)\to F(m_2+_cm_3)\xrightarrow{x_1\times1}F(m_1)\times F(m_2+_cm_3)\xrightarrow{\phi}F(m_1\times(m_2+_cm_3))$}&\scalebox{.8}{$\to F(m_1+_b(m_2+_cm_3))$}
\end{align*}
%%% Kenny: I think we should call this \alpha something else like \kappa, since it's only a component used in building the associator for horizontal composition (which we are calling \alpha), not the entire thing on its own.
\node (A) at (3.75,0) {$\scriptstyle{\cong \tau_a}$};
\node (D) at (3,0) {$\scriptstyle{\one}$};
\node (E) at (6.25,0.5) {$\scriptstyle{F(m_1 + m_2) \times F(m_3)}$};
\node () at (7.625,0) {$\cong$};
%\node () at (8.75,0) {$\cong$};
\node (E') at (6.25,-0.5) {$\scriptstyle{F(m_1) \times F(m_2 + m_3)}$};
\node (B) at (8,0.5) {$\scriptstyle{F((m_1 + m_2) + m_3)}$};
\node (B') at (8,-0.5) {$\scriptstyle{F(m_1 + (m_2 + m_3))}$};
\node (C) at (9.5,0.5) {$\scriptstyle{F((m_1+_b m_2)+m_3)}$};
\node (C') at (9.5,-0.5) {$\scriptstyle{F(m_1 + (m_2+_c m_3))}$};
\node (F) at (11,0.5) {$\scriptstyle{F((m_1+_b m_2)+_c m_3)}$};
\node (F') at (11,-0.5) {$\scriptstyle{F(m_1+_b(m_2+_c m_3))}$};
\node (G) at (4.25,0.5) {$\scriptstyle{(F(m_1) \times F(m_2)) \times F(m_3)}$};
\node (G') at (4.25,-0.5) {$\scriptstyle{F(m_1) \times (F(m_2) \times F(m_3))}$};
\node (M) at (3.25,0.3) {$\scriptstyle{(x_1 \times x_2) \times x_3}$};
\node (M') at (3.25,-0.3) {$\scriptstyle{x_1 \times (x_2 \times x_3)}$};
(C) edge node [above] {$F(\psi)$} (F)
(C') edge node [above] {$F(\psi)$} (F')
(F) edge node [right] {$F(\kappa)$} (F')
(E) edge node [above] {$\phi_{m_1+m_2,m_3}$} (B)
(E') edge node [above] {$\phi_{m_1,m_2+m_3}$} (B')
%(C) edge node  [fill=white] {$F(h_1 +_g h_2)$} (C')
%(B) edge node [fill=white] {$F(h_1 + h_2)$} (B')
(G) edge node [above] {$\phi_{m_1,m_2} \times 1$} (E)
(G') edge node [above] {$1 \times \phi_{m_2,m_3}$} (E')
(G) edge node [fill=white] {$a$} (G')
(D) edge node [left,above] {$$} (G)
(D) edge node [below] {$$} (G');
%(E) edge node [fill=white] {$F(h_1) \times F(h_2)$} (E');
where $a$ is the associator of $\Cat$ and $\kappa$ is the canonical isomorphism between two colimits of the same diagram in $\A$. Similarly the right and left unitors can be constructed, and all data satisfy the axioms of a double category -- essentially breaking down to those of the cospan double category of $\A$ and axioms of the lax monoidal pseudofunctor $F$. For full details of these arguments, see \cite[Theorem~4.1.1]{CourserThesis}.
 The definition of double category may be found in Definition \ref{defn:double_category}. To show that the pseudofunctor $F$ in the theorem statement induces a double category $F\lCsp$, we must first define the category of objects $F\lCsp_0$, the category of arrows $F\lCsp_1$, as well as the four structure functors which tie them together: the source and target functors $S,T \colon F\lCsp_1 \to F\lCsp_0$, the unit functor $U \colon F\lCsp_0 \to F\lCsp_1$ and the composition functor $\odot$ which says how to compose two composable horizontal 1-cells or two composable 2-morphisms horizontally: $$\odot \colon F\lCsp_1 \times_{F\lCsp_0} F\lCsp_1 \to F\lCsp_1.$$

@@

Let $\A$ be a category with finite colimits and $F \maps \A \to \Cat$ a symmetric lax monoidal pseudofunctor. Then the double category $F\lCsp$ of Theorem \ref{thm:decorated_cospans} is symmetric monoidal, where

@@

\begin{proof}
The definition of symmetric monoidal double category may be found in Definition \ref{defn:symmetric_monoidal_double_category}. We must show that both the category of objects $F\lCsp_0$ and the category of arrows $F\lCsp_1$ are symmetric monoidal categories, show that the monoidal unit of $F\lCsp_0$ is preserved by the unit functor $U$, show that the source and target functors $S,T \colon F\lCsp_1 \to F\lCsp_0$ are strict symmetric monoidal functors which preserve the associativity and unit constraints, define the globular isomorphisms $\chi$ and $\mu$ that relate how composition and units interact with the tensor product, and check that the required axioms are satisfied.

First we note that the category of objects $F\lCsp_0=\A$ is symmetric monoidal under chosen binary coproducts in $\A$ and the left and right unitors, associators and braidings are given as natural maps. Next we show that $F\lCsp_1$ is symmetric monoidal. The category of arrows $F\lCsp_1$ has:
\begin{enumerate}
\item{objects as $F$-decorated cospans, which are pairs:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
and}
\item{morphisms as maps of $F$-decorated cospans, which are cospans in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node [above]{$i'$} (D)
(C') edge node [above]{$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a decoration morphism $\tau \maps F(h)(x) \to x'$.
}
\end{enumerate}
First we define the tensor product and monoidal unit of $F\lCsp_1$. Given two objects $M_1$ and $M_2$ of $F\lCsp_1$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$m_1$};
\node (C) at (2,0) {$b_1$};
\node (D) at (1,-0.5) {$x_1 \in F(m_1)$};
\node (E) at (3,0) {$a_2$};
\node (F) at (4,0) {$m_2$};
\node (G) at (5,0) {$b_2$};
\node (H) at (4,-0.5) {$x_2 \in F(m_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(E) edge node[above]{$i_2$} (F)
(G) edge node[above]{$o_2$} (F);
\end{tikzpicture}
\]
their tensor product $M_1 \otimes M_2$ is given by taking the coproducts of the cospans in $\A$
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1+a_2$};
\node (B) at (1.5,0) {$m_1+m_2$};
\node (C) at (3,0) {$b_1+b_2$};
\node (D) at (4.55,0) {$x_1+x_2 \in F(m_1+m_2)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+i_2$} (B)
(C) edge node[above]{$o_1+o_2$} (B);
\end{tikzpicture}
\]
where the decoration on the apex is obtained using the natural transformation of the symmetric lax monoidal pseudofunctor $F$: $$x_1+x_2 \mapseqq 1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{x_1 \times x_2} F(m_1) \times F(m_2) \xrightarrow{\phi_{m_1,m_2}} F(m_1+m_2).$$The monoidal unit $0_{F\lCsp_1}$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$0$};
\node (B) at (1,0) {$0$};
\node (C) at (2,0) {$0$};
\node (D) at (3,0) {$\bot_0 = \phi \in F(0)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$!$} (B)
(C) edge node[above]{$!$} (B);
\end{tikzpicture}
\]
where $0$ is the monoidal unit of $\A$ and $\phi \maps 1 \to F(0)$ is the morphism which is part of the structure of the symmetric lax monoidal pseudofunctor $F \maps \A \to \Cat$. It is also clear that $U(0)=0_{F\lCsp_1}$.

Next we define the associator and left unitor of $F\lCsp_1$, with the right unitor being similar to the left. Given three objects $M_1, M_2$ and $M_3$ of $F\lCsp_1$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,0) {$m_1$};
\node (C) at (2,0) {$b_1$};
\node (D) at (1,-0.5) {$x_1 \in F(m_1)$};
\node (E) at (3,0) {$a_2$};
\node (F) at (4,0) {$m_2$};
\node (G) at (5,0) {$b_2$};
\node (H) at (4,-0.5) {$x_2 \in F(m_2)$};
\node (I) at (6,0) {$a_3$};
\node (J) at (7,0) {$m_3$};
\node (K) at (8,0) {$b_3$};
\node (L) at (7,-0.5) {$x_3 \in F(m_3)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1$} (B)
(C) edge node[above]{$o_1$} (B)
(E) edge node[above]{$i_2$} (F)
(G) edge node[above]{$o_2$} (F)
(I) edge node[above]{$i_3$} (J)
(K) edge node[above]{$o_3$} (J);
\end{tikzpicture}
\]
tensoring the first two and then the third results in $(M_1 \otimes M_2) \otimes M_3$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$(a_1+a_2)+a_3$};
\node (B) at (2.5,0){$(m_1+m_2)+m_3$};
\node (C) at (5,0) {$(b_1+b_2)+b_3$};
\node (D) at (2.5,-0.5) {$(x_1+x_2)+x_3 \in F((m_1+m_2)+m_3)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$(i_1+i_2)+i_3$} (B)
(C) edge node[above]{$(o_1+o_2)+o_3$} (B);
\end{tikzpicture}
\]
where $(x_1+x_2)+x_3 \maps 1 \to F((m_1+m_2)+m_3)$ is given by: $$1 \xrightarrow{(x_1 \times x_2) \times x_3} (F(m_1) \times F(m_2)) \times F(m_3) \xrightarrow{\phi_{m_1,m_2} \times 1} F(m_1+m_2) \times F(m_3) \xrightarrow{\phi_{m_1+m_2,m_3}} F((m_1+m_2)+m_3).$$ The other parenthesization $M_1 \otimes (M_2 \otimes M_3)$ is given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1+(a_2+a_3)$};
\node (B) at (2.5,0) {$m_1+(m_2+m_3)$};
\node (C) at (5,0) {$b_1+(b_2+b_3)$};
\node (D) at (2.5,-0.5) {$x_1+(x_2+x_3) \in F(m_1+(m_2+m_3))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i_1+(i_2+i_3)$} (B)
(C) edge node[above]{$o_1+(o_2+o_3)$} (B);
\end{tikzpicture}
\]
where $x_1+(x_2+x_3) \maps 1 \to F(m_1+(m_2+m_3))$ is given by: $$1 \xrightarrow{x_1 \times (x_2 \times x_3)} F(m_1) \times (F(m_2) \times F(m_3)) \xrightarrow{1 \times \phi_{m_2,m_3}} F(m_1) \times F(m_2+m_3) \xrightarrow{\phi_{m_1,m_2+m_3}} F(m_1+(m_2+m_3)).$$
Denoting the associator of $(\A,+,0)$ by $\alpha$, the associator of $F\lCsp_1$ is then a map of cospans in $\A$ from $(M_1 \otimes M_2) \otimes M_3$ to $M_1 \otimes (M_2 \otimes M_3)$ given by:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$(a_1+a_2)+a_3$};
\node (A') at (0,-0.5) {$a_1+(a_2+a_3)$};
\node (B) at (2.25,0.5) {$(m_1+m_2)+m_3$};
\node (C) at (4.5,0.5) {$(b_1+b_2)+b_3$};
\node (C') at (4.5,-0.5) {$b_1+(b_2+b_3)$};
\node (D) at (2.25,-0.5) {$m_1+(m_2+m_3)$};
\node (E) at (7,0.5) {$(x_1+x_2)+x_3 \in F((m_1+m_2)+m_3)$};
\node (F) at (7,-0.5) {$x_1+(x_2+x_3) \in F(m_1+(m_2+m_3))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$(i_1+i_2)+i_3$} (B)
(C) edge node[above]{$(o_1+o_2)+o_3$} (B)
(A) edge node[left]{$\alpha$} (A')
(C) edge node[right]{$\alpha$} (C')
(A') edge node [above]{$i_1+(i_2+i_3)$} (D)
(C') edge node [above]{$o_1+(o_2+o_3)$} (D)
(B) edge node [left] {$\alpha$} (D);
\end{tikzpicture}
\]
together with the decoration isomorphism $\tau_\alpha \maps F(\alpha)((x_1+x_2)+x_3) \to x_1+(x_2+x_3)$.

For the left unitor, we first compute $0_{F\lCsp_1} \otimes M$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$0$};
\node (B) at (1,0) {$0$};
\node (C) at (2,0) {$0$};
\node (D) at (1,-0.5) {$\bot_0 \in F(0)$};
\node (E) at (3,0) {$a$};
\node (F) at (4,0) {$m$};
\node (G) at (5,0) {$b$};
\node (H) at (4,-0.5) {$x \in F(m)$};
\node (I) at (2.5,0) {$\otimes$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$!$} (B)
(C) edge node[above]{$!$} (B)
(E) edge node[above]{$i$} (F)
(G) edge node[above]{$o$} (F);
\end{tikzpicture}
\]
which results in:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$0+a$};
\node (B) at (1,0) {$0+m$};
\node (C) at (2,0) {$0+b$};
\node (D) at (3.5,0) {$\bot_0 + x \in F(0+m)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$!+i$} (B)
(C) edge node[above]{$!+o$} (B);
\end{tikzpicture}
\]
where $\bot_0 + x \in F(0+m)$ is given by $$1 \xrightarrow{\lambda^{-1}} 1 \times 1 \xrightarrow{\bot_0 \times x} F(0) \times F(m) \xrightarrow{\phi_{0,m}} F(0+m).$$Denoting the left unitor of $(\A,+,0)$ by $\ell$, the left unitor of $F\lCsp_1$ is then an isomorphism in $F\lCsp_1$ given by the following map of cospans in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$0+a$};
\node (A') at (0,-0.5) {$a$};
\node (B) at (1,0.5) {$0+m$};
\node (C) at (2,0.5) {$0+b$};
\node (C') at (2,-0.5) {$b$};
\node (D) at (1,-0.5) {$m$};
\node (E) at (3.5,0.5) {$\bot_0 + x \in F(0+m)$};
\node (F) at (3.5,-0.5) {$x \in F(m)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$!+i$} (B)
(C) edge node[above]{$!+o$} (B)
(A) edge node[left]{$\ell$} (A')
(C) edge node[right]{$\ell$} (C')
(A') edge node [above]{$i$} (D)
(C') edge node [above]{$o$} (D)
(B) edge node [left] {$\ell$} (D);
\end{tikzpicture}
\]
together with the decoration isomorphism $\tau_{\ell} \maps F(\ell)(\bot_0 + x) \to x$. The right unitor is similar, and one can check that the associator and left and right unitors together satisfy the pentagon and triangle identities of a monoidal category. See the second author's thesis \cite{CourserThesis} for more details.
